---
title: "R Functions: Question 6"
author: "Juan Manuel (PID: A69034758)"
format: pdf
---
```{r}
# Can you improve this analysis code?
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Exploring the data frame structure... 

```{r}
s1 <- read.pdb("4AKE")
```

```{r}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.chainA
```

```{r}
s1.b <- s1.chainA$atom$b
s1.b
```

By exploring the type of functions and data from the above code, we can see that they run the functions three times. To simplify this code, we will need to define a function that iterates through the three PBD strcutures. 
For this, we first need to built a vector that contains all the PBD IDs. 
```{r}
library(bio3d)
pdb_ids <- c("4AKE", "1AKE", "1E4Y") #kinase with drug 4AKE, #kinase no drug 1AKE, kinase with drug 1E4Y
```

After this, we need to built a function that integrates the four main functions in the original code. That is: reading the data, trimming the data, extracts the B-factors for the alpha carbon atoms in chain A of the PDB file, and plot them. 

```{r}
B_factors_plots <- function(pdb_indivual_id) {
  pdb_data <- read.pdb(pdb_indivual_id) #reading data
  chain_data <- trim.pdb(pdb_data, chain = "A", elety = "CA") #triming data
  b_factors <- chain_data$atom$b #getting b factors
  plotb3(b_factors, sse=chain_data, typ="l", ylab="B-factor", main=pdb_indivual_id) #plotting data
}
```
As we need to iterate this function to every single item from pdb_ids, we are gonna be using the function sapply(). 
```{r}
final_plots <- sapply(pdb_ids, B_factors_plots) #using the function in the pdb_ids list. 
```

And then we obtained the same output as the original code but optimized. 


